\documentclass[11pt,letterpaper]{article}

\usepackage[margin=1in]{geometry}
\usepackage{termcal}
\usepackage{enumitem}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{color}
\usepackage{multirow}
\usepackage{multicol}

\newcommand{\todo}[1]{\textcolor{red}{TODO: #1}}

\title{16-715: Advanced Robot Dynamics and Simulation}
\author{Fall 2021}
\date{}

\begin{document}

\maketitle

\section*{Course Description}

This course explores the fundamental mathematics behind modeling the physics of robots, as well as state-of-the-art algorithms for robot simulation. We will review classical topics like Lagrangian mechanics and Hamiltons Principle of Least Action, as well as modern computational methods like discrete mechanics and fast linear-time algorithms for dynamics simulation. A particular focus of the course will be rigorous treatments of 3D rotations and non-smooth contact interactions (impacts and friction) that are so prevalent in robotics applications. We will use numerous case studies to explore these topics, including quadrotors, fixed-wing aircraft, wheeled vehicles, quadrupeds, humanoids, and manipulators. Homework assignments will focus on practical implementation of algorithms and a course project will encourage students to apply simulation methods to their own research.

\medskip
\noindent
\textbf{Prerequisites:} Strong linear algebra skills, experience with a high-level programming language like Python, MATLAB, or Julia, and basic familiarity with ordinary differential equations.

\section*{Instructors}

\begin{center}
\begin{tabular}{l l}
	Prof. Zac Manchester & \textbf{Email:} \href{mailto:zacm@cmu.edu}{zacm@cmu.edu} \\
	TA: Kevin Tracy & \textbf{Email:} \href{mailto:ktracy.cmu.edu}{ktracy@andrew.cmu.edu}
\end{tabular}
\end{center}

\section*{Logistics}

\begin{itemize}
	\item Lectures will be held Tuesdays and Thursdays 10:10--11:30 AM Eastern time in NSH 1305.
	\item Office hours will be \todo{based on survey}.
	\item Homework assignments will be due by 11:59 PM Eastern time on Wednesdays. Two weeks will be given to complete each assignment.
	\item GitHub will be used to distribute and collect assignments.
	\item Slack will be used for general discussion and Q\&A outside of class and office hours.
	\item There will be no exams. Instead, each student will complete a project on a topic of their choice.
\end{itemize}

\section*{Learning Objectives}
By the end of this course, students should be able to:
\begin{enumerate}
	\item Derive differential equations for simple mechanical systems using classical Newton-Euler and Lagrangian techniques
	\item Model more complex industrial, wheeled, legged, aerial, underwater, and space robotic systems using modern computational techniques
	\item Simulate environmental contact interactions like impacts and Coulomb friction
	\item Understand the trade-offs and limitations of different dynamics formulations and simulation techniques
	\item Build high-performance simulation tools that can be deployed in machine learning and control design applications
	
\end{enumerate}

\section*{Learning Resources}

There is no textbook required for this course. Video recordings of lectures and lecture notes will be posted online. Additional references for further reading will be provided with each lecture.

\section*{Homework}

Homework will be posted every 2 weeks and students will be given at least one full week to complete assignments. All homework will be distributed and collected using GitHub. Solutions and grades will be returned within one week of homework due dates.

\section*{Grading}

Grading will be based on:
\begin{itemize}
	\item 50\% Project
	\item 40\% Homeworks
	\item 10\% Participation
\end{itemize}
Attendance during lectures is not required to earn a full participation grade. Students can also participate through any combination of office hours, Slack discussions, project presentations, and by offering constructive feedback about the course to the instructors.


\section*{Course Policies}

\textbf{Late Homework:} Students are allowed a budget of 6 late days for turning in homework with no penalty throughout the semester. They may be used together on one assignment, or separately on two assignments. Beyond these six days, no other late homework will be accepted.

\medskip
\noindent
\textbf{Accommodations for Students with Disabilities:} If you have a disability and are registered with the Office of Disability Resources, I encourage you to use their online system to notify me of your accommodations and discuss your needs with me as early in the semester as possible. I will work with you to ensure that accommodations are provided as appropriate. If you suspect that you may have a disability and would benefit from accommodations but are not yet registered with the Office of Disability Resources, I encourage you to contact them at \href{mailto:access@andrew.cmu.edu}{access@andrew.cmu.edu}.

\medskip
\noindent
\textbf{Statement of Support for Students' Health \& Well-Being:} Take care of yourself. Do your best to maintain a healthy lifestyle this semester by eating well, exercising, avoiding drugs and alcohol, getting enough sleep, and taking some time to relax. This will help you achieve your goals and cope with stress.

\medskip
\noindent
If you or anyone you know experiences any academic stress, difficult life events, or feelings like anxiety or depression, we strongly encourage you to seek support. Counseling and Psychological Services (CaPS) is here to help: call 412-268-2922 and visit \href{http://www.cmu.edu/counseling}{http://www.cmu.edu/counseling}. Consider reaching out to a friend, faculty, or family member you trust for help getting connected to the support that can help.

\medskip
\noindent
\textit{If you or someone you know is feeling suicidal or in danger of self-harm, call someone immediately, day or night:}

\textit{CaPS: 412-268-2922}

\textit{Re:solve Crisis Network: 888-796-8226}

\medskip
\noindent
\textit{If the situation is life threatening, call the police:}

\textit{On campus: CMU Police: 412-268-2323}

\textit{Off campus: 911}

\medskip
\noindent
If you have questions about this or your coursework, please let me know. Thank you, and have a great semester.


\section*{Tentative Schedule}

\begin{tabular}{c|c|c|c}
	Week & Dates & Topics & Assignments \\
	\hline
	\multirow{2}{*}{1} & Aug 31 & Course Overview \& Review of Newtonian Mechanics Concepts \\
	 & Sep 2 & Particles, Pendulums, and Orbits &  \\
	\hline
	\multirow{2}{*}{2} & Sep 7 & Energy and Stability & HW 1 Out \\
	 & Sep 9 & Numerical Solution of ODEs and Runge-Kutta Methods &  \\
	\hline
	\multirow{2}{*}{3} & Sep 14 &  \textcolor{red}{No Class} &  \\
	 & Sep 16 & Rigid Bodies, Euler's Equation, and Lie Groups &  \\
	\hline
	\multirow{2}{*}{4} & Sep 21 & Quaternions and Numerical Simulation & HW 1 Due \\
	 & Sep 23 & Quadrotors, Airplanes, and Spacecraft & HW 2 Out \\
	\hline
	\multirow{2}{*}{5} & Sep 28 & Constrained Optimization Pt. 1 & \\
	 & Sep 30 & The Least-Action Principle &  \\
	\hline
	\multirow{2}{*}{6}  & Oct 5 & Coordinates, Constraints, and Manifolds &   HW2 Due \\
	 & Oct 7 & D'Alembert, Virtual Work, and Generalized Forces & HW3 Out \\
	\hline
	\multirow{2}{*}{7}  & Oct 12 & Lagrangian Mechanics and Manipulators & \\
	 & Oct 14 & \textcolor{red}{No Class} & \\
	\hline
	\multirow{2}{*}{8}  & Oct 19 & Momentum, Duality, and the Hamiltonian & \\
	 & Oct 21 & Lie Algebras and Euler's Equation from a Lagrangian & \\
	\hline
	\multirow{2}{*}{9}  & Oct 26 & Discrete Mechanics and Variational Integrators & HW3 Due \\
	 & Oct 28 & Constrained Optimization Pt. 2 &   \\
	\hline
	\multirow{2}{*}{10}  & Nov 2 & Impacts as Inequality Constraints &  HW4 Out \\
	 & Nov 4 & Coulomb Friction and The Maximum Dissipation Principle & \\
	 \hline
	\multirow{2}{*}{11}  & Nov 9 & Legged Robots and Hybrid Systems &  \\
	 & Nov 11 & Sparsity and Linear-Time Algorithms &   \\
	 \hline
	\multirow{2}{*}{12}  & Nov 16 & Case Studies &  HW4 Due \\
	 & Nov 18 & Case Studies &   \\
	 \hline
	\multirow{2}{*}{13}  & Nov 23 & Case Studies &  \\
	 & Nov 25 & \textcolor{red}{No Class} &   \\
	 \hline
	\multirow{2}{*}{14}  & Nov 30 & Project Presentations &  \\
	 & Dec 2 & Project Presentations &   \\
\end{tabular}

\section*{Project Guidelines}

Students should work in groups of 1--4 to complete a substantial final project. The goal is for students to apply the coarse content to their own research. Project proposals will be solicited in late September and topics will be selected in consultation with the instructors.

\medskip
\noindent
Project grades will be based on a short presentation given during the last week of class and a final report submitted via \href{https://forms.gle/j1xhW13DuvcYLf6j7}{Google drive} by May 18 \href{https://time.is/Anywhere_on_Earth}{Anywhere on Earth}. Reports should be written in the form of a 6 page (plus references) ICRA or IROS conference paper using the standard \href{https://www.ieee.org/conferences/publishing/templates.html}{two-column IEEE format}. Sections should include an abstract, introduction and/or background to motivate your problem, 2--3 main technical sections on your contributions, conclusions, and references. Grading will be based on the following criteria:
\newline
\newline
\begin{tabular}{|c|l|}
\hline
10\% & Class presentation \\
\hline
10\% & Adherence to IEEE formatting and length requirements \\
\hline
10\% & Innovation \& Creativity: Is what you did new/cool/interesting? Convince me. \\
\hline
30\% & Clarity of presentation: Can I understand what you did from your writing + plots? \\
\hline
40\% & Technical correctness: Are your results reasonable? Is your code correct? \\
\hline	
\end{tabular}


\end{document}
